---
title: "Drawing up plays with ggulti"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Drawing up plays with ggulti}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6,
  fig.path = "."
)
```

## Getting Started

Install ggulti from github with:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("swebb1/ggulti")
```

Load the ggulti package:

```{r setup, message=FALSE, warning=FALSE}
library(ggulti)
```

## Create a pitch

Use the `ggpitch()` function to create a pitch. The `type` argument specifies the type of pitch to create. The `endzone_fill` argument specifies the color of the endzone.

```{r, message=FALSE, warning=FALSE}
pitch = ggpitch(type="half_attack",endzone_fill = "#AA4422", ymin = 50)
```

Pitch types

-   "full" = Full pitch (default)

-   "half_attack" = Half a pitch showing attacking endzone (up)

-   "half_defend" = Half a pitch showing defending endzone (down)

-   "blank" = Lined out area. Useful for drills etc.

-   "void" = No pitch background.

```{r, message=FALSE, warning=FALSE}
ggpitch()
```

```{r, message=FALSE, warning=FALSE}
ggpitch(type = "half_attack", fill = "lightgreen", alpha=1, ymin = 50)
```

```{r, message=FALSE, warning=FALSE}
ggpitch(type = "half_defend", endzone_fill = "black", ymax = 50)
```

```{r, message=FALSE, warning=FALSE}
ggpitch(type = "blank", ymax = 50, xmax = 30)
```

```{r, message=FALSE, warning=FALSE}
ggpitch(type = "void")
```

The `ggpitch()` function returns a *ggplot* object that can be customised further.

## Pitch objects

Pitch objects are plotted as points and are typically players, discs and cones. Use the `pitch_object()` function to create object and their positions.

Let's create a list of three offense players attacking the endzone with the `pitch_object()` function:

```{r, message=FALSE, warning=FALSE}
players <- list(
  pitch_object(label = "H1", show = T, x = c(5), y = c(75), alpha = 0.8, frame = 1:3),
  pitch_object(label = "H2", show = T, x = c(20), y = c(75), alpha = 0.5, frame = 1:3),
  pitch_object(label = "C1", show = T, x = c(18,14,5), y = c(83,83,98), alpha = 0.8, frame = 1:3)
)
```

In the code above, we have two stationary handlers (H1 and H2) and one cutter (C1) making a cut along the front of the endzone and to the back corner.

### What are frames?

Frames are used to animate objects. In the code above, the `frame` argument specifies the frames that the object is visible. In this case, the objects are visible in frames 1, 2 and 3. H1 and H2 are stationary objects so only have single x and y coordinates. C1 is a moving object so has multiple x and y coordinates for each frame. The frame argument is optional and defaults to 1 (e.g. a single image).

We can plot out our players on the pitch we created using the `plot_play()` function. The default is to plot frame 1.

```{r, message=FALSE, warning=FALSE}
plot_play(pitch = pitch, object_list = players)
```

We can also choose which frame to plot:

```{r, message=FALSE, warning=FALSE}
plot_play(pitch = pitch, object_list = players, static_frame = 3)
```

Or we can plot all frames and objects which might be useful when creating our plots:

```{r, message=FALSE, warning=FALSE}
plot_play(pitch = pitch, object_list = players,show_all = T)
```

### Other object types

By default, the object type is set to "Offense". The following object types are recognised

-   "Offense" = Green circle

-   "Defense" = Blue circle

-   "Coach" = Red circle

-   "Cone" = Orange triangle

-   "Disc" = Small grey circle


```{r, message=FALSE, warning=FALSE}
random <- list(
  pitch_object(label = "Queue-here", show = T, object = "Cone", x = c(5), y = c(75), alpha = 0.8, frame = 1),
  pitch_object(label = "Pep", show = T, object = "Coach", x = c(20), y = c(75), alpha = 0.5, frame = 1),
  pitch_object(label = "disc", show = F, object = "Disc", x = c(18), y = c(83), alpha = 0.8, frame = 1),
  pitch_object(label = "D1", show = F, object = "Defense", x = c(18), y = c(55), alpha = 0.8, frame = 1)
)

plot_play(pitch = pitch, object_list = random)
```

### Custom objects

You can also add your own custom objects by specifying a different `object` argument.

```{r, message=FALSE, warning=FALSE}
custom <- list(
  pitch_object(label = "Ref", show = T, object = "Referee", x = c(18), y = c(55), alpha = 0.8, frame = 1)
)

plot_play(pitch = pitch, object_list = custom)
```

By default the object will present as an offense object. You can change the default options (colour, size, shape) in the `plot_play()` function:

```{r, message=FALSE, warning=FALSE}
plot_play(pitch = pitch, object_list = custom, default_obj_col = "black")
```

Alternatively, you can use the `pv` argument to add a new list of plotting values. The `ggulti_plot_values` object gives access to default values:

```{r, message=FALSE, warning=FALSE}
pv <- ggulti_plot_values
pv$obj_cols <- c(pv$obj_cols,"Referee" = "black")
pv$obj_shapes <- c(pv$obj_shapes,"Referee" = 18)
pv$obj_sizes <- c(pv$obj_sizes,"Referee" = 2)

plot_play(pitch = pitch, object_list = custom, pv = pv)
```

By customising objects, you can specifiy specific teams, genders or even psoitions e.g.

```{r, message=FALSE, warning=FALSE}
zone_positions <- c("Cup" = "blue", "Wing" = "red", "Mid" = "purple", "Deep" = "green")
pv <- ggulti_plot_values
pv$obj_cols <- c(ggulti_plot_values$obj_cols, zone_positions)

zone <- list(
  pitch_object(label = "C1", show=T, object = "Cup", x=c(14), y=c(25)),
  pitch_object(label = "C2", show=T, object = "Cup", x=c(18), y=c(28)),
  pitch_object(label = "C3", show=T, object = "Cup", x=c(22), y=c(25)),
  pitch_object(label = "W1", show=T, object = "Wing", x=c(10), y=c(40)),
  pitch_object(label = "W2", show=T, object = "Wing", x=c(27), y=c(40)),
  pitch_object(label = "Mid", show=T, object = "Mid", x=c(18), y=c(38), label_pos = "up"),
  pitch_object(label = "Deep", show=T, object = "Deep", x=c(18), y=c(60))
)

plot_play(pitch = ggpitch(endzone_fill="darkgrey"),object_list = zone,pv=pv)

```

## Add defenders

Let's return to our original set of three offense players attacking the endzone. We would like to add a defender to each player.

We can add defenders using the `pitch_object()` function and specifying the object type as "Defense". We can also use the `defender()` function to add defenders to existing players. Defenders are objects that follow another object through frames. You can set their position relative to the object using the `xjust` and `yjust` arguments.

Defenders inherit defaults from their player (frame,alpha, show). The default for `xjust` and `yjust` is -1. This means the defender will be placed 1 unit to the left and 1 unit below the object, or slightly under on the backhand side. You can use single values for the offsets or a value for each frame.

```{r message=FALSE, warning=FALSE}
## Add defenders to players
defenders <- list(
  defender(player = players[[1]],label = "D1", show = T, xjust = c(2), yjust = c(2)),
  defender(player = players[[2]],label = "D2", show = T, xjust = c(-3), yjust = c(2)),
  defender(player = players[[3]],label = "D3", show = T, xjust = c(-3,-3,3), yjust = c(-1,-1,-5))
)
```

Let's now combine our players and defenders into a single list:

```{r message=FALSE, warning=FALSE}
## Join players and defenders
player_list <- c(players,defenders)
plot_play(pitch = pitch,object_list = player_list, keep_arrows = T, static_frame = 1)
```

## Pitch arrows

Pitch arrows can be added in a similar way to objects using the `pitch_arrow()` function. However, arrows are typically used to represent player movements or throws between two players.

### Throws

We can create a list of throws using the `throw()` function. This function requires an object_list containing objects (e.g. players) to throw `to` and `from`. The start and end points of a throw are determined by the involved players positions in a specific `throwing_frame`.

The `to` argument can also be set to "space" for a throw into empty space. In this case `space_x` and `space_y` are required to specify the target location.

You can also specify an `arrow_shape` to show straight or curved throws:

-   "straight" = Straight arrow (default)

-   "fhrc or bhio" = Right curved arrow

-   "bhrc or fhio" = Left curved arrow


```{r message=FALSE, warning=FALSE}
## Add throws between players
throws <- list(
  throw(object_list = player_list, label = "T1", show = F,from = "H1",to = "C1",frame = 1:2, throw_frame = 3, arrow_shape = "bhrc")
)
```

### Player paths

You can also add arrows to represent object movements (e.g. player cuts) using the `object_paths()` function. This function requires an object_list containing objects (e.g. players) to plot paths for. The `objects` argument specifies the object types to plot paths for. By default this is set to "Offense" to only plot paths for Offense player objects.

```{r message=FALSE, warning=FALSE}
paths <-  object_paths(player_list) ## Add arrows in front of running players
```

Now we can add our throws and paths to create a list of arrows:

```{r message=FALSE, warning=FALSE}
## Create arrow list
arrow_list = list(
  paths,
  throws
)
```

## Adding a disc

We can add discs to our graphics with the `pitch_object()` function using `object="Disc"` and specifying the coordinates and frames. Alternatively we can use the `disc()` function to create a disc that follows a specific throw. This function requires a `pitch_arrow` object to trace. The `release_frame` argument specifies the frame where the disc is thrown.

```{r message=FALSE, warning=FALSE}
## Add a disc
disc_list = list(
  disc(throw = throws[[1]],label = "disc",show = F,frame = 1:3,release_frame = 3)
)
```

The disc should be added to our list of pitch objects:

```{r message=FALSE, warning=FALSE}
## Create object list
object_list = c(player_list,disc_list)
```

## Plotting the play

We now have objects and arrows representing players, defenders, cuts, throws and a disc. We can plot the play using the `plot_play()` function:

```{r message=FALSE, warning=FALSE}
plot_play(pitch = pitch,object_list = object_list,arrow_list = arrow_list,keep_arrows = T,static_frame = 1)
```

```{r message=FALSE, warning=FALSE}
plot_play(pitch = pitch,object_list = object_list,arrow_list = arrow_list,keep_arrows = T,static_frame = 2)
```

```{r message=FALSE, warning=FALSE}
plot_play(pitch = pitch,object_list = object_list,arrow_list = arrow_list,keep_arrows = T,static_frame = 3)
```

## Animating the play

The big advantage of setting up plays with frames is that we can create animations. We can animate the play by setting the `animate = T` argument. Animations use `gganimate` with transition_states. The speed of the animation can be controlled with the `transition_time` and `state_time` arguments.

```{r message=FALSE, warning=FALSE}
plot_play(pitch = pitch,object_list = object_list,arrow_list = arrow_list,keep_arrows = T,animate = T,transition_length = 1,state_length = 0.5)
```

## Further customisation

Because the `plot_play` function returns a ggplot object, we can further customise the plot using ggplot2 functions. For example, we can add a title and change the theme:

```{r message=FALSE, warning=FALSE}
library(ggplot2)

plot_play(pitch = pitch,object_list = object_list,arrow_list = arrow_list,keep_arrows = T,static_frame = 1) +
  ggtitle("Back of the zone!") +
  geom_label(aes(x = 18, y = 94, label = "Score!"), size = 7, color = "black") +
  theme_classic()
```

For more control over animations, you can save the `plot_play()` output and run [`gganimate`](https://gganimate.com/index.html) yourself.

```{r message=FALSE, warning=FALSE}
library(gganimate)
library(gifski)

p <- plot_play(pitch = pitch,object_list = object_list,arrow_list = arrow_list,show_all = T)

ap <- p +
transition_states(states = factor(frame, levels = 1:3), transition_length = 0.5, state_length = 0.5, wrap=F) +
  shadow_mark(size = 0, colour = 'grey') +
  enter_fade()

animate(ap, renderer = gifski_renderer())
```
